VM-Centric Snapshot Deduplication\\ for Cloud Data Backup

### 2015-05-07 snapshot_system/mass_eval.tex
change for msst

\section{Evaluation}
\label{sect:evaluation}
We have implemented and evaluated a prototype of our VC scheme on a Linux cluster of machines with
8-core 3.1Ghz AMD FX-8120 and 16 GB RAM. 
Our implementation is based on the Alibaba cloud platform~\cite{Aliyun,WeiZhangIEEE}
and the underlying DFS uses  QFS with default replication degree 3 while the PDS replication degree is 6.
Our evaluation objective is to
study deduplication efficiency and resource usage of VC,  
and assess its processing time and backup throughput, and  
the benefit in fault tolerance. 
We will compare VC with a VO approach  using stateless routing with binning (SRB) 
based on~\cite{Dong2011,extreme_binning09}.
SRB executes distributed deduplication by routing data chunks to cluster machines~\cite{Dong2011}
using  a min-hash function discussed in \cite{extreme_binning09}. Once a data chunk is routed to
a machine, the chunk is compared with the fingerprint index within this machine locally based 
on~\cite{extreme_binning09}.

%{\bf Settings}.
We have performed a trace-driven study based on a production dataset
from Alibaba Aliyun's cloud platform~\cite{Aliyun} 
with about 2500 VMs, running on 100 physical machines. 
Each machine in the production environment is more powerful than our evaluation cluster
and has 16 cores with 48GB memory.
Each machine hosts up to 25 VMs and each VM keeps 10 automatically-generated snapshots in the storage 
system while a user may instruct extra snapshots to be saved.
Each VM has about 40GB of storage  data  on average
including OS and user data disk while the size distribution is highly skewed and the maximum to
the average ratio is close to 40.
Each physical machine deals with about 10TB of snapshot data and each 1TB data represents one snapshot version
of 25 VMs.  Thus the total amount of raw snapshot shot in 100 physical machines
is about 1,000TB.
The VMs of the sampled data set use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, win2008 and win2003. 
%The backup of VM snapshots is required to complete within few  hours every night.
The daily snapshot change rate is about 2-3\% on average.
%All data are divided into 2 MB fix-sized segments and 
The fingerprint for variable-sized chunks is computed using their SHA-1 hash~\cite{similar94,rabin81}. 
%Each segment is divided into 
%variable-sized content chunks with an average size of 4KB.
%Popularity of data blocks are collected through global counting and 
%and top popular ones ($\delta=2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:store}.

@@

\comments{
%old table, delete 4%
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline 
&
     \multicolumn{4}{|c|}{VC}  & VO \\ \hline
&
     2\%PDS  &  No PDS &2\%PDS     & 4\%PDS & SRB  \\ 
&     no-simi&   &&  &    \\ \hline
Dedup &     94.31\%& 93.02\%  & 96.01\% & 96.58\% & 97.86\% \\ \hline
Mem(MB) &     108&  46 & 118  & 190 & 2400 \\ \hline
    \end{tabular}
}
\begin{table}[htb]
\centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline 
&
     \multicolumn{3}{|c|}{VC}  & VO \\ \hline
&       No PDS & No PDS &  2\%PDS     &   SRB  \\ 
&       no simi &simi   &  simi    &\\ \hline
Dedup & 75.86\%    &87.91\%  & 96.01\%   & 97.86\% \\ 
Efficiency & &           &       &                 \\    \hline 
Memory &     N/A&   10MB & 118MB  & 2.4GB \\ 
per-machine & &     &   &    \\ \hline
    \end{tabular}
    \caption{Deduplication efficiency and per-machine memory usage for different VC settings and
        SRB.}
    \label{tab:efficiency-comparison}
\end{table}

@@

We explain why similarity-guided local search provides such an improvement as follows.
%Without this (i.e. when we only 
There are VMs in which data  segments are moved to another location on disk, for example 
when a file is rewritten
rather than modified in place,  
and a dirty-bit or offset-only based detection would not be able to detect such a movement.
We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently.
%Column 4 of Table~\ref{tab:efficiency-comparison} shows deduplication efficiency of VC which only uses
%the same offset in the parent segment to look for duplicates  but does not look for other potentially
%similar segments. 
Compared an approach in ~\cite{WeiZhangIEEE} which maintains additional offset dirty bits within each segment,
 local similarity-guided search adds additional  deduplication efficiency about 3\%,
which is a saving of 300GB per machine.
%from 94\% to over 96\%.
%That is one significant improvement compared to the work in 
%which uses the parent segment at the same offset to detect duplicates instead of similarity-guided search.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.

@@

Comparing the experiment results of the sampled index method reported in~\cite{Guo2011}, 
our scheme achieves a 85K:1 ratio between raw snapshot data and memory usage 
with 96\% deduplication efficiency
while the sampled index method achieves 20K:1 (25GB memory per 500TB raw data)
with 97\% deduplication efficiency.
Thus our scheme is more memory efficient with a good tradeoff for deduplication efficiency.
If we change $\sigma=4\%$, the deduplication efficiency of VC increases to 96.58\% while
memory usage per machine increases to about 220MB. 
Notice also that the sampled index method is designed for single-machine deduplication.
%and it is not easy to extend to a distributed cluster architecture.

@@

{\bf Throughput of software layers without I/O throttling.}
Table~\ref{tab:throughput} shows the  average throughput of software layers
when I/O throttling is not applied to control usage. The
%machine when all machine nodes execute one or multiple MV backup tasks.
``Backup'' column is the throughput  of the backup service  per machine.
``Snapshot store" is the  write throughput of the snapshot store layer and the significant reduction from this
column to  ``Backup" column is caused by deduplication.
Only non-duplicate chunks trigger a snapshot store write.
Column ``QFS'' is the write request traffic to the underlying file system after compression.
For example, with 148 MB/s write traffic to the snapshot store, QFS write traffic is about 35.3 MB/s
after compression.  However, the underlying disk storage traffic will be three times greater with replication.
The result shows that the backup service can deliver up to 5.46 GB/s 
per machine without I/O restriction
under 6 concurrent backup tasks. 
For our dataset, each version of total snapshots has abut 1TB per machine for 25 VMs and  thus  each machine
would complete the backup in about 3.05 minutes.
With a higher disk storage bandwidth available, the above backup
throughput would be higher. 
In comparison, the sampled index method in~\cite{Guo2011} achieves about 5.5GB per second
with 16 concurrent backup tasks and 6GB per second with
more tasks.  Thus the per-machine throughput of our scheme is reasonable.
%With 50MB/second controlled I/O bandwidth, each machine can deliver 171MB/second with 1 backup task and can complete the backup of 25 VMs per machine in less than 1.31 hours.

>>>>

\section{Evaluation}
\label{sect:evaluation}
We have implemented and evaluated a prototype of our VC scheme on a Linux cluster of machines with
8-core 3.1Ghz AMD FX-8120 and 16 GB RAM. 
Our implementation is based on the Alibaba cloud platform~\cite{Aliyun,WeiZhangIEEE}
and the underlying DFS uses  QFS with default replication degree 3 while the PDS replication degree is 6.
Our evaluation objective is to
study deduplication efficiency and resource usage of VC,  
and assess its processing time and backup throughput, and  
the benefit in fault tolerance. 
We will compare VC with a VO approach  using stateless routing with binning (SRB) 
based on the previous work~\cite{Dong2011,extreme_binning09}.
SRB executes distributed deduplication by routing data chunks to cluster machines~\cite{Dong2011}
using  a min-hash function~\cite{extreme_binning09}. Once a data chunk is routed to
a machine, the chunk is compared with the fingerprint index within this machine 
locally~\cite{extreme_binning09}.
%MSST-CHANGE-BEGIN
While this VO approach is designed for general deduplication and data backup,
we choose this VO approach as a baseline to compare because the other
approaches would have  a similar level of computing resource usage.
%MSST-CHANGE-END

%{\bf Settings}.
We have performed a trace-driven study based on a production dataset
from Alibaba Aliyun's cloud platform~\cite{Aliyun} 
with about 2500 VMs, running on 100 physical machines. 
Each machine in the production environment is more powerful than our evaluation cluster
and has 16 cores with 48GB memory.
Each machine hosts up to 25 VMs and each VM keeps 10 automatically-generated snapshots in the storage 
system while a user may instruct extra snapshots to be saved.
Each VM has about 40GB of storage  data  on average
including OS and user data disk while the size distribution is highly skewed and the maximum to
the average ratio is close to 40.
Each physical machine deals with about 10TB of snapshot data and each 1TB data represents one snapshot version
of 25 VMs.  Thus the total amount of raw snapshot in 100 physical machines
is about 1,000TB.
The VMs of the sampled data set use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, win2008 and win2003. 
%The backup of VM snapshots is required to complete within few  hours every night.
%MSST-CHANGE-BEGIN
Each machine hosts a variety of applications running these operating systems including web services and 
database management. The amount of daily data update is not significant and
the daily snapshot change rate is about 1-3\% on average.
%MSST-CHANGE-END
%All data are divided into 2 MB fix-sized segments and 
The fingerprint for variable-sized chunks is computed using their SHA-1 hash~\cite{similar94,rabin81}. 
Note that variable sizing is not required in our scheme and we just follow the setting in the previous work.
%Each segment is divided into 
%variable-sized content chunks with an average size of 4KB.
%Popularity of data blocks are collected through global counting and 
%and top popular ones ($\delta=2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:store}.

@@

\comments{
%old table, delete 4%
    \begin{tabular}{|c|c|c|c|c|c|}
    \hline 
&
     \multicolumn{4}{|c|}{VC}  & VO \\ \hline
&
     2\%PDS  &  No PDS &2\%PDS     & 4\%PDS & SRB  \\ 
&     no-simil&   &&  &    \\ \hline
Dedup &     94.31\%& 93.02\%  & 96.01\% & 96.58\% & 97.86\% \\ \hline
Mem(MB) &     108&  46 & 118  & 190 & 2400 \\ \hline
    \end{tabular}
}
\begin{table}[htb]
\centering
    \begin{tabular}{|c|c|c|c|c|}
    \hline 
&
     \multicolumn{3}{|c|}{VC}  & VO \\ \hline
&       No PDS & No PDS &  2\%PDS     &   SRB  \\ 
&       no simil &simil   &  simil    &\\ \hline
Dedup & 75.86\%    &87.91\%  & 96.01\%   & 97.86\% \\ 
Efficiency & &           &       &                 \\    \hline 
Memory &     N/A&   10MB & 118MB  & 2.4GB \\ 
per-machine & &     &   &    \\ \hline
    \end{tabular}
    \caption{Deduplication efficiency and per-machine memory usage for different VC settings and
        SRB.}
    \label{tab:efficiency-comparison}
\end{table}

@@

We explain why similarity-guided local search provides such an improvement as follows.
%Without this (i.e. when we only 
There are VMs in which data  segments are moved to another location on disk, for example 
when a file is rewritten
rather than modified in place,  
and a dirty-bit or offset-only based detection would not be able to detect such a movement.
We have found that in
approximately 1/3 of the VMs in our dataset this movement happens frequently.
%Column 4 of Table~\ref{tab:efficiency-comparison} shows deduplication efficiency of VC which only uses
%the same offset in the parent segment to look for duplicates  but does not look for other potentially
%similar segments. 
Compared a previous  approach~\cite{WeiZhangIEEE} which maintains additional offset dirty bits within each segment,
 local similarity-guided search adds additional  deduplication efficiency about 3\%,
which is a saving of 300GB per machine.
%from 94\% to over 96\%.
%That is one significant improvement compared to the work in 
%which uses the parent segment at the same offset to detect duplicates instead of similarity-guided search.
% In general, our experiments show that
% dirty-bit detection at the segment level can reduce the data size to about 24.14\% of original data, 
% which leads about a 75.86\% reduction.
% Similarity-guided local search can further reduce the data size
% to about 12.05\% of original size, namely it delivers an additional 12.09\% reduction.
% The PDS-based global duplicate detection with $\sigma=2\% $
% can reduce the size further to 8.6\% of original size, namely it 
% delivers an additional 3.45\% reduction.

@@

Comparing the experiment results of the sampled index method~\cite{Guo2011}, 
our scheme achieves a 85K:1 ratio between raw snapshot data and memory usage 
with 96\% deduplication efficiency
while the sampled index method achieves 20K:1 (25GB memory per 500TB raw data)
with 97\% deduplication efficiency.
Thus our scheme is more memory efficient with a good tradeoff for deduplication efficiency.
If we change $\sigma=4\%$, the deduplication efficiency of VC increases to 96.58\% while
memory usage per machine increases to about 220MB. 
Notice also that the sampled index method is designed for single-machine deduplication.
%and it is not easy to extend to a distributed cluster architecture.

@@

{\bf Throughput of software layers without I/O throttling.}
Table~\ref{tab:throughput} shows the  average throughput of software layers
when I/O throttling is not applied to control usage. The
%machine when all machine nodes execute one or multiple MV backup tasks.
``Backup'' column is the throughput  of the backup service  per machine.
``Snapshot store" is the  write throughput of the snapshot store layer and the significant reduction from this
column to  ``Backup" column is caused by deduplication.
Only non-duplicate chunks trigger a snapshot store write.
Column ``QFS'' is the write request traffic to the underlying file system after compression.
For example, with 148 MB/s write traffic to the snapshot store, QFS write traffic is about 35.3 MB/s
after compression.  However, the underlying disk storage traffic will be three times greater with replication.
The result shows that the backup service can deliver up to 5.46 GB/s 
per machine without I/O restriction
under 6 concurrent backup tasks. 
For our dataset, each version of total snapshots has abut 1TB per machine for 25 VMs and  thus  each machine
would complete the backup in about 3.05 minutes.
With a higher disk storage bandwidth available, the above backup
throughput would be higher. 
In comparison, the sampled index method~\cite{Guo2011} achieves about 5.5GB per second
with 16 concurrent backup tasks and 6GB per second with
more tasks.  Thus the per-machine throughput of our scheme is reasonable.
%With 50MB/second controlled I/O bandwidth, each machine can deliver 171MB/second with 1 backup task and can complete the backup of 25 VMs per machine in less than 1.31 hours.

### 2015-05-07 snapshot_system/mass_intro.tex
re
change for msst

%Without deduplication,
Source-side deduplication (e.g. ~\cite{Symantec2009})
 eliminates duplicates before backup data is transmitted
to a secondary storage, which saves network bandwidth significantly; however its resource
usage can impact other co-located cloud services. 
It is memory-intensive to compare a large number of fingerprints and identify
duplicates, even with optimization or approximation techniques developed~\cite{Guo2011,Dong2011,extreme_binning09,FuFast2015}. 
\comments{
When deleting unused snapshots, a grouped mark-and-sweep approach~\cite{Guo2011}  
is effective while it still carries a significant cost, especially in a distributed
setting. 
} Another side effect of deduplication is the possible loss of failure
resilience~\cite{Reliability06}.  
%Separate files share the same physical copies of blocks that are
%logically duplicated among them.  
If a shared block is lost, all files that
share that block are affected.  
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out raw and undeduplicated backup data wastes a huge amount
of network bandwidth that would otherwise be available to user VMs. 
\comments{
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).
}

Our work  considers a backup service with source-side deduplication for converged cloud architectures
that is co-located with user VMs, sharing the cloud compute, network, and
storage resource with them.  
This paper extends our preliminary simulation study  in ~\cite{WeiZhangIEEE} to 
restrict the scope of cross-VM deduplication and simplify the deduplication process
by separating popular data chunks and presents the design and implementation of 
a VM-centric backup scheme.  
We term this approach {\em VM-centric} because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
Our work focuses on a tradeoff that allows the sharing of only ``popular'' data blocks 
across virtual machines while using localized deduplication within each VM
to achieve both storage savings and fault isolation.
Another issue addressed is fault  isolation in a content sharing environment caused by deduplication.
One consequence of aggressive deduplication is the possible loss of failure
resilience.  If a shared block is lost, all files that
share that block are affected~\cite{Reliability06} and  our work is motivated by this
to develop VM-centric fault isolation techniques.
\comments{
Since the underlying storage file system block size is typically bigger than the average data chunk size
used for deduplication,  packaging
content from different VMs into the same file block  creates a ``false'' sharing that 
affects fault isolation;
}

The main contribution of this paper is
1) a design and implementation of a backup service 
to integrate the VM-centric techniques while minimizing the resource impact on other 
collocated cloud services;
2) a VM-centric file block management to improve snapshot availability by
separating popular chunk replication and
packaging data chunks from the same VM into a file system block as much as possible; 
3) an analysis on the deduplication efficiency by using popular items
and also on snapshot availability improved by the above fault isolation design. 
The integrated optimization includes  similarity-guided local deduplication for
each VM to make the overall deduplication efficiency more competitive 
to a traditional deduplication scheme.
Our  experimental study compares  the VM-centric approach with the previous 
VM-oblivious approaches in terms of deduplication efficiency, resource usage,  and snapshot availability.
%Thus we improve  to avoid such VM-oblivious packaging.
%our approach uses a deduplication data chunk size that is smaller than a
%single file block and

>>>>

%Without deduplication,
Source-side deduplication~\cite{Vrable2009,Tan2010,Fu2014,Symantec2009}
eliminates duplicates before backup data is transmitted
to a secondary storage, which saves network bandwidth significantly; however its resource
usage can impact other co-located cloud services. 
It is memory-intensive to compare a large number of fingerprints and identify
duplicates, even with optimization or approximation techniques developed~\cite{Guo2011,Dong2011,extreme_binning09,FuFast2015}. 
\comments{
When deleting unused snapshots, a grouped mark-and-sweep approach~\cite{Guo2011}  
is effective while it still carries a significant cost, especially in a distributed
setting. 
} Another side effect of deduplication is the possible loss of failure
resilience~\cite{Reliability06}.  
%Separate files share the same physical copies of blocks that are
%logically duplicated among them.  
If a shared block is lost, all files that
share that block are affected.  
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out raw and undeduplicated backup data wastes a huge amount
of network bandwidth that would otherwise be available to user VMs. 
\comments{
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).
}

Our work  considers a backup service with source-side deduplication for converged cloud architectures
that is co-located with user VMs, sharing the cloud compute, network, and
storage resource with them.  
This paper extends our preliminary simulation study~\cite{WeiZhangIEEE} to 
restrict the scope of cross-VM deduplication and simplify the deduplication process
by separating popular data chunks and presents the design and implementation of 
a VM-centric backup scheme.  
We term this approach {\em VM-centric} because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
Our work focuses on a tradeoff that allows the sharing of only ``popular'' data blocks 
across virtual machines while using localized deduplication within each VM
to achieve both storage savings and fault isolation.
Another issue addressed is fault  isolation in a content sharing environment caused by deduplication.
One consequence of aggressive deduplication is the possible loss of failure
resilience.  If a shared block is lost, all files that
share that block are affected~\cite{Reliability06} and  our work is motivated by this
to develop VM-centric fault isolation techniques.
\comments{
Since the underlying storage file system block size is typically bigger than the average data chunk size
used for deduplication,  packaging
content from different VMs into the same file block  creates a ``false'' sharing that 
affects fault isolation;
}

The main contribution of this paper is
1) a design and implementation of a backup service 
to integrate the VM-centric techniques while minimizing the resource impact on other 
collocated cloud services;
2) a VM-centric file block management to improve snapshot availability by
separating popular chunk replication and
packaging data chunks from the same VM into a file system block as much as possible; 
3) an analysis on the deduplication efficiency by using popular items
and also on snapshot availability improved by the above fault isolation design. 
The analysis provides insights and justification on the proposed VM-centric strategies.
The integrated optimization includes  similarity-guided local deduplication for
each VM to make the overall deduplication efficiency more competitive 
to a traditional deduplication scheme.
Our  experimental study compares  the VM-centric approach with the previous 
VM-oblivious approaches in terms of deduplication efficiency, resource usage,  and snapshot availability.
Compared to our earlier study~\cite{WeiZhangIEEE} which achieves 92\% of what the full deduplication
can do, our integrated system delivers over  96\% in a tested dataset from Alibaba cloud. 
%Thus we improve  to avoid such VM-oblivious packaging.
%our approach uses a deduplication data chunk size that is smaller than a
%single file block and
